Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-08 vesicle_code/docs/adhesion/adhesion.tex
thrusday night 1st revisions by yny

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet.  
When the long axis of the vesicle doublet is aligned with the far-field diverging flow, the converging flow
pushes vesicles towards each other and the repulsive interaction is essential to keep the two vesicles at the separation distance.
In this configuration the extensional flow is pushing the vesicles together. With the stagnation point placed at the center of the doublet, this
configuration appears to be quite stable for low extensional flow rate.

However, when the long axis of vesicle doublet is parallel
to the far-field converging flow, the diverging flow may overcome the vesicle adhesion to
split the two vesicles apart. Thus in the following simulation experiments we place the vesicle doublet so that the flat contact region is
aligned with the axis of converging flow. In this configuration the flow is pulling the vesicles away from each and 
is expected to be less stable than the orthogonal configuration where the fluid flow is constantly pushing vesicles towards each other. 
It is also reasonable to expect the vesicle doublet to rotate around the stagnation so the flow is pushing them together rather than pulling them apart.

At low extensional
rates, we expect the vesicle doublet to stay at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

\begin{figure}[htp]
  \begin{minipage}{0.53\textwidth}
    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalPhaseDiagram} {\em Left}: The
  definition of the inclination angle.  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the diverging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the diverging direction.  The square
  mark is a stagnation point and the round point is located at the
  center of the vesicle. {\em Right}: A phase diagram of the long-time
  behavior of a fluid trap formed by two adhering vesicles in an
  extensional flow.  The Hamaker constant is $\mathcal{H}=0.7$ and the
  adhesion length scale is $\delta = 0.4$.  For small extensional rates,
  the doublet structure is not broken (blue marks), while at large
  extensional rates, the doublet is broken (red marks).  The size of the
  round blue marks indicates the final inclination angle with the larger
  dots rotating less ($\theta=\pi$) and the smaller dots rotating more
  ($\theta = \pi/2$).  A square mark is used for simulations that did
  not result in an equilibrium inclination angle in the given time
  horizon.}
\end{figure}

>>>>

\begin{figure}[htp]
  \begin{minipage}{0.53\textwidth}
    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
  \end{minipage}
   \caption{\label{fig:InclinationAngle} The
  definition of the inclination angle.  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the diverging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the diverging direction.  The square
  mark is a stagnation point and the round point is located at the
  center of the vesicle.}
\end{figure}

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet, which has a flat contact region as shown in \S~\ref{sec:qflow}. 
Based on how the contact region aligns with the direction of divergent flow, 
we can define the inclination angle of the vesicle doublet as illustrated in Figure~\ref{fig:InclinationAngle}.
%

When $\theta=\pi$ (the long axis of the vesicle doublet is aligned with the far-field diverging flow), the converging flow
pushes vesicles towards each other and the short-range repulsive force is essential to keep the two vesicles at the separation distance.
With the fluid flow pushing the two vesicles towards the stagnation point dynamically placed at the doublet center, 
this configuration is expected to be more stable than the configuration in the middle of Figure~\ref{fig:InclinationAngle} where $\theta=\pi/2$ 
(the long axis of vesicle doublet is parallel to the converging direction).  As shown in the illustration, 
the diverging flow is pulling the vesicles away from each other when $\theta=\pi/2$. 
Thus it is possible that the diverging flow may overcome the adhesion and
break up the doublet. It is also reasonable to expect the vesicle doublet to rotate around the stagnation so the flow is pushing them together rather than pulling them apart.

In the following simulation experiments we place the vesicle doublet with $\theta=\pi/2$ so that the flat contact region is
orthogonal to the diverging flow that is pulling vesicles away from each other.
%In this configuration the flow is pulling the vesicles away from each other, and 
%is expected to be less stable than the $\theta=\pi$ configuration where the fluid flow is constantly pushing vesicles towards each other. 
%
At low extensional
rates, we expect the vesicle doublet to stay at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

\begin{figure}[htp]
%  \begin{minipage}{0.53\textwidth}
%    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
%    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
%    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
%  \end{minipage}
%  \hfill
  \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalPhaseDiagram}  A phase diagram of the long-time
  behavior of a fluid trap formed by two adhering vesicles in an
  extensional flow.  The Hamaker constant is $\mathcal{H}=0.7$ and the
  adhesion length scale is $\delta = 0.4$.  For small extensional rates,
  the doublet structure is not broken (blue marks), while at large
  extensional rates, the doublet is broken (red marks).  The size of the
  round blue marks indicates the final inclination angle with the larger
  dots rotating less ($\theta=\pi$) and the smaller dots rotating more
  ($\theta = \pi/2$).  A square mark is used for simulations that did
  not result in an equilibrium inclination angle in the given time
  horizon.}
\end{figure}
